---
title: "Class1 survey analysis"
author: "Vishnuvardhan"
date: "2024-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##### 1. Import class one survey data from our Github site (https://github.com/kijohnson/ADA-2024) (The dataset is called 'Class 1 Survey Fall 2024_di.csv'), calling the R dataframe that you create *C1survey*.
``````{r}
# reading the C1survey dataset from Github
pacman::p_load(dplyr, readr)
Class1survey <- read_csv("https://raw.githubusercontent.com/kijohnson/ADA-2024/main/Class%201%20(Getting%20started%20and%20RR%20I)/Lab/Class%201%20Survey%20Fall%202024_di.csv")
View(Class1survey)
```
##### 2. Determine: **a.** the number of observations (i.e. the number of people who filled out the survey) and **b.** the number of variables in the dataframe.
``````{r}
# a and b determine number of observations and variables in C1ass1survey dataset
dim(Class1survey)
# there are 29 observations and 27 variables
str(Class1survey)
```
##### 3. **a.** Rename the column variables to something shorter and that is descriptive of what the variable is about (for example *like_dogs* for the 'Do you like dogs?' question variable) and **b.** Display that they are renamed.
```{r}
#using names function to rename all the 27 variables in sequence
names(Class1survey)[1:27]<-c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three","public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city", "highest_educ_level")

#displaying the renamed variables
colnames(Class1survey)
```
##### 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the *C1survey* dataset.
```{r}
table(sapply(Class1survey, class)) #using table and sapply functions 23 character and 4 integer
summary(Class1survey)
```
##### 5. **a.** Using code, check the *bday* and *bmonth* variables for any unusual or missing values. If you find any, **b.** describe what you will do with the unusual values in a comment before or immediately following the code that does it, and **c.** after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics
```{r}
#a.looking at values for bday and bmonth
Class1survey$bday
Class1survey$bmonth

#b.recoding bday with non numeric values

#bday recoding method
class(Class1survey$bday) # checking that it is coded as character class
Class1survey$bday<-ifelse(Class1survey$bday == "May 21-report 21", "21", Class1survey$bday)
Class1survey$bday<-ifelse(Class1survey$bday == "March 31st", "31", Class1survey$bday)
Class1survey$bday # check values 

# convert to numeric for statistics
Class1survey$bday<-as.numeric(Class1survey$bday)
table(Class1survey$bday, useNA = "always") # checking values 

# bmonth recoding
class(Class1survey$bmonth)

Class1survey$bmonth[Class1survey$bmonth=="May 21-report 5"]<-"5"
Class1survey$bmonth[Class1survey$bmonth=="February"]<-"2"
Class1survey$bmonth[Class1survey$bmonth=="September"]<-"9"
Class1survey$bmonth[Class1survey$bmonth=="July"]<-"7"
Class1survey$bmonth[Class1survey$bmonth=="March"]<-"3"
Class1survey$bmonth # check values 

#convert to numeric for statistics
Class1survey$bmonth<-as.numeric(Class1survey$bmonth)
Class1survey$bmonth # check values 

#c.calculating medians for bday and bmonth 
median(Class1survey$bday, na.rm=TRUE) # 14 (biased because of 0, which is not a value) 
median(Class1survey$bmonth, na.rm=TRUE) # 7
```

##### 6. **a.** Create a new variable called *bseason* that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). **b**. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct **c**. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code. 
```{r}
# a.
# using dplyr
Class1survey <- Class1survey %>%
  mutate(bseason=case_when(bmonth %in% c(12,1,2) ~"Winter",
                           bmonth %in% c(3,4,5) ~"Spring",
                           bmonth %in% c(6,7,8) ~"Summer",
                           bmonth %in% c(9,10,11) ~"Fall"))
table(Class1survey$bmonth, Class1survey$bseason) # check variable

Class1survey$bseason_f<-factor(Class1survey$bseason, levels=c("Winter", "Spring", "Summer", "Fall"))


# b. 
table(Class1survey$bmonth, Class1survey$bseason)
table(Class1survey$bmonth, Class1survey$bseason_f)

# c.
tab<-addmargins(table(Class1survey$bmonth, Class1survey$bseason, useNA = "always"), 1)
tab
tab<-addmargins(table(Class1survey$bmonth, Class1survey$bseason_f, useNA = "always"), 1)
tab

# Births by season: Fall = 4 Winter = 9 Spring = 8 Summer = 8
```
##### 7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #'s).

#FROM THE GIVEN DATASET, WHAT IS THE AVERAGE LEVEL OF CODING COMFORT?
```{r}
View(Class1survey)
mean(Class1survey$coding_comfort, na.rm=TRUE) 
#2.75
#USING mean function
```

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
